% Preface
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Preface} % Abstract name
\addtocontents{toc}{\protect\vspace{\beforebibskip}} % Place the bibliography slightly below the rest of the document content in the table of contents
\addcontentsline{toc}{chapter}{\tocEntry{Preface}}

Finding general principles underlying brain function has been appealing to scientists.
Indeed, in some branches of science like physics and chemistry (and to some degree biology) a general theory often can capture the essence of a wide range of phenomena.
Whether we can find such principles in neuroscience, and [assuming they do exist]
what those principles are, are important questions.
Abstracting the brain as a complex system is one of the perspectives that
may help us answer this question.

While it is commonly accepted that the brain is a (or even \emph{the}) prominent example of a complex system, 
the far reaching implications of this fact are still arguably overlooked in our approaches to neuroscientific questions.
% 
One of the reasons for the lack of attention could be the apparent difference in foci of investigations in these two fields --- neuroscience and complex systems.
This thesis is an effort toward providing a bridge between systems neuroscience and complex systems
by harnessing systems neuroscience tools \& notions for building empirical approaches toward the brain as a complex system.


Perhaps, in the spirit of \emph{searching for principles},
we should abstract and approach the brain as a complex \emph{adaptive} system as the more complete perspective (rather than just a complex system).
In the end, the brain, even the most ``complex system'', need to survive in the environment.
Indeed, in the field of \emph{complex adaptive systems}, the intention is understanding very similar questions in nature.
As an outlook, we also touch on some research directions pertaining to the adaptivity of the brain as well.

\endgroup			

\vfill

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../phdThesis_csb"
%%% End:
