\begin{description}
\item[Title:]
  Decoding the contents of consciousness from prefrontal ensembles
\item[Authors:]
  Vishal Kapoor$^*$,
  Abhilash Dwarakanath$^*$,
  Shervin Safavi,
  Joachim Werner,
  Michel Besserve,
  Theofanis I. Panagiotaropoulos,
  Nikos K. Logothetis
  ($^*$ indicate equal contributions) 
\item[Status:]
  Accepted for publication in Nature Communication (preprint is available online, see \citet{kapoorDecodingInternallyGenerated2022})
\item[Presentation at scientific meetings:]
  FFRM 2015 \cite{antoniouPerceptualModulationPupillary2015a},
  SfN 2018 \cite{panagiotaropoulosModulationNeuralDischarges2018a},
  FENS 2018 \cite{kapoorSpikingActivityPrefrontal2018},
  ASSC 2019 \cite{kapoorNeuronalDischargesPrefrontal2019}
\item[Author contributions: ]
  V.K., A.D. and T.I.P. designed the study. V.K., A.D. and S.S. trained animals. V.K. and A.D.
  performed experiments and collected data, with occasional help from S.S. V.K. and A.D.
  analyzed the data. S.S. contributed to spike sorting and selectivity analysis of control
  experiments. M.B. contributed to the decoding analysis. V.K. prepared and arranged the figures
  in the final format. S.S. provided the MATLAB generated version of the figures displayed in
  figure 3D, S12, S13 and S14 A. T.I.P. and N.K.L. supervised the study. N.K.L. and J.W.
  contributed unpublished reagents/analytical tools. N.K.L. provided the support to the group. V.K.
  and T.I.P. wrote the original manuscript draft. All authors participated in discussion and
  interpretation of the results and editing the manuscript.
\end{description}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../phdThesis_csb"
%%% End:
